title: "map gene count to annotation"
author: "YNYK"
date: "`r Sys.Date()`"
output:
  html_document:
  number_sections: no
toc: yes
toc_depth: 6
toc_float: true
self_contained: true
thumbnails: true
lightbox: true
gallery: true
use_bookdown: false
highlight: haddock
---

## Load package
```{r}
library(dplyr)
library(tibble)
library(tidyr)
library(foreach)
library(doParallel)
library(arrow)
library(rhdf5)
```

## Load data
```{r}
eggNog_df <- read.delim2("Genecatalog/annotations/eggNog.tsv",sep = "\t",header=2)[c(-1),]
eggNog_df<-eggNog_df %>% dplyr::filter(!grepl("#query",eggNog_df$Query))
rownames(eggNog_df)<-eggNog_df$Query


##gene name

gene <- read_parquet("Genecatalog/counts/gene_coverage_stats.parquet")
colnames(gene)[1]<-"gene_name"
gene_name<-stringr::word(gene$gene_name,1,1,sep=stringr::fixed(" "))


filename = "Genecatalog/counts/median_coverage.h5"
Nmapped_count <- h5read(filename, "data")
attributes= h5readAttributes(filename, "data")
colnames(Nmapped_count) <- attributes$sample_names %>% make.names()
rownames(Nmapped_count)<-gene_name


# Nmapped_count <- read.delim2("Genecatalog/counts/Nmapped_reads.tsv",sep = "\t",header=1) %>% column_to_rownames("X.ID")
annot_count<-merge(eggNog_df,Nmapped_count,by=0)
dim(annot_count)
```


## KEGG_Module 

```{r}
count_colname<-c("Query","KEGG_Module",colnames(Nmapped_count) )
KEGG_Module <- annot_count[,count_colname]
KEGG_Module <- KEGG_Module[KEGG_Module$KEGG_Module != "-",]
KEGG_Module <- KEGG_Module[,-1]


KEGG_Module_extended <- c()
file.remove("Genecatalog/annotations/KEGG_Module_count.tsv")


write.table(colnames(KEGG_Module) %>% t(),file = "Genecatalog/annotations/KEGG_Module_count.tsv",append = T,sep="\t",quote = F,row.names = F,col.names = F)

for (i in 1 : dim(KEGG_Module)[1]){
  # i=1
  id=KEGG_Module[i,]
  if (grepl(",",id$KEGG_Module)){
    KEGG_Modules <- strsplit(id$KEGG_Module,split = ",")[[1]]
    for(n in 1:length(KEGG_Modules)){
      # n=1
      tmp <- data.frame(KEGG_Modules[n],id[-1] )
      colnames(tmp)[1] <- "KEGG_Module"
      KEGG_Module_extended <- bind_rows(KEGG_Module_extended,tmp )
    }
  }else{
    KEGG_Module_extended <- bind_rows(KEGG_Module_extended,id)
  }
  
  if(i %% 500 == 0){
    
    #print(i)
    
    write.table(KEGG_Module_extended,file = "Genecatalog/annotations/KEGG_Module_count.tsv",append = T,sep="\t",quote = F,row.names = F,col.names = F)
    KEGG_Module_extended <- c()
    
  }
}


# read table
KEGG_Module_c <- read.delim2("Genecatalog/annotations/KEGG_Module_count.tsv")
KEGG_Module_c$KEGG_Module <- gsub("KEGG_Module:","",KEGG_Module_c$KEGG_Module)

KEGG_Module_c2 <- KEGG_Module_c %>% arrange(KEGG_Module) %>% 
  group_by(KEGG_Module) %>%
  summarise_all(.funs = sum)
KEGG_Module_c2 <- KEGG_Module_c2 %>% data.frame()
write.csv(KEGG_Module_c2,file = "Genecatalog/annotations/KEGG_Module_count_grouped.csv",row.names = F,quote = F)
```




### KEGG_Module filtered
```{r}
kegg_modules <- read.delim2("genomes/annotations/dram/kegg_modules.tsv",header = T)  
kegg_modules$step_coverage =kegg_modules$step_coverage %>% as.numeric()
kegg_modules=kegg_modules%>% dplyr::filter(step_coverage>=0.8)
kegg_modules=kegg_modules%>% dplyr::select(module,module_name) %>% dplyr::distinct(module,.keep_all = T)
# kegg_modules$module_name <- stringr::word(kegg_modules$module_name,1,1,stringr::fixed(","))
# kegg_modules$name_id<- paste(kegg_modules$module,"_",kegg_modules$module_name)
# kegg_modules=kegg_modules %>% dplyr::select(-module_name)

KEGG_Module_c2.modules=merge(kegg_modules,KEGG_Module_c2,by.x = "module",by.y ="KEGG_Module" ,all.x = T)
write.csv(KEGG_Module_c2.modules,file = "Genecatalog/annotations/KEGG_Module_count_grouped_step_coverage0.8.csv",row.names = F,quote = T)
```


## KO table (one gene to many KO)
```{r,message=F}
count_colname<-c("Query","KO",colnames(Nmapped_count) )
KO <- annot_count[,count_colname]
KO <- KO[KO$KO != "-",]
KO <- KO[,-1]


KO_extended <- c()
file.remove("Genecatalog/annotations/KO_count.tsv")


write.table(colnames(KO) %>% t(),file = "Genecatalog/annotations/KO_count.tsv",append = T,sep="\t",quote = F,row.names = F,col.names = F)

for (i in 1 : dim(KO)[1]){
  id=KO[i,]
  if (grepl(",",id$KO)){
    kos <- strsplit(id$KO,split = ",")[[1]]
    for(n in 1:length(kos)){
      tmp <- data.frame(kos[n],id[-1] )
      colnames(tmp)[1] <- "KO"
      KO_extended <- bind_rows(KO_extended,tmp )
    }
  }else{
  KO_extended <- bind_rows(KO_extended,id)
  }
  
  if(i %% 500 == 0){
    
    #print(i)
    
    write.table(KO_extended,file = "Genecatalog/annotations/KO_count.tsv",append = T,sep="\t",quote = F,row.names = F,col.names = F)
    KO_extended <- c()
    
    }
}


# read table
KO_c <- read.delim2("Genecatalog/annotations/KO_count.tsv")
KO_c$KO <- gsub("ko:","",KO_c$KO)

KO_c2 <- KO_c %>% arrange(KO) %>% 
  group_by(KO) %>%
  summarise_all(.funs = sum)

write.csv(KO_c2,file = "Genecatalog/annotations/KO_count_grouped.csv",row.names = F,quote = F)

unique(KO_c$KO) %>% length()

```




## CAZy (one gene to many )
```{r}
count_colname<-c("Query","CAZy",colnames(Nmapped_count) )
ca <- annot_count[,count_colname]
ca <- ca[ca$CAZy != "-",]
ca <- ca[,-1]


ca_extended <- c()
file.remove("Genecatalog/annotations/CAZy_count.tsv")


write.table(colnames(ca) %>% t(),file = "Genecatalog/annotations/CAZy_count.tsv",append = T,sep="\t",quote = F,row.names = F,col.names = F)

for (i in 1 : dim(ca)[1]){
  id=ca[i,]
  if (grepl(",",id$CAZy)){
    cas <- strsplit(id$CAZy,split = ",")[[1]]
    for(n in 1:length(cas)){
      tmp <- data.frame(cas[n],id[-1] )
      colnames(tmp)[1] <- "CAZy"
      ca_extended <- bind_rows(ca_extended,tmp )
    }
  }else{
  ca_extended <- bind_rows(ca_extended,id)
  }
  
  if(i %% 500 == 0){
    
    #print(i)
    
    write.table(ca_extended,file = "Genecatalog/annotations/CAZy_count.tsv",append = T,sep="\t",quote = F,row.names = F,col.names = F)
    ca_extended <- c()
    
    }
}


# read table
ca_c <- read.delim2("Genecatalog/annotations/CAZy_count.tsv")


ca_c2 <- ca_c %>% arrange(CAZy) %>% 
  group_by(CAZy) %>%
  summarise_all(.funs = sum)

write.csv(ca_c2,file = "Genecatalog/annotations/CAZy_count_grouped.csv",row.names = F,quote = F)

unique(ca_c$CAZy) %>% length()


```


## EC (one gene to many )
```{r}
count_colname<-c("Query","EC",colnames(Nmapped_count) )
EC <- annot_count[,count_colname]
EC <- EC[EC$EC != "-",]
EC <- EC[,-1]


EC_extended <- c()
file.remove("Genecatalog/annotations/EC_count.tsv")


write.table(colnames(EC) %>% t(),file = "Genecatalog/annotations/EC_count.tsv",append = T,sep="\t",quote = F,row.names = F,col.names = F)

for (i in 1 : dim(EC)[1]){
  id=EC[i,]
  if (grepl(",",id$EC)){
    ECs <- strsplit(id$EC,split = ",")[[1]]
    for(n in 1:length(ECs)){
      tmp <- data.frame(ECs[n],id[-1] )
      colnames(tmp)[1] <- "EC"
      EC_extended <- bind_rows(EC_extended,tmp )
    }
  }else{
  EC_extended <- bind_rows(EC_extended,id)
  }
  
  if(i %% 500 == 0){
    
    #print(i)
    
    write.table(EC_extended,file = "Genecatalog/annotations/EC_count.tsv",append = T,sep="\t",quote = F,row.names = F,col.names = F)
    EC_extended <- c()
    
    }
}


# read table
EC_c <- read.delim2("Genecatalog/annotations/EC_count.tsv")
#EC_c$EC <- gsub("EC:","",EC_c$EC)

EC_c2 <- EC_c %>% arrange(EC) %>% 
  group_by(EC) %>%
  summarise_all(.funs = sum)

write.csv(EC_c2,file = "Genecatalog/annotations/EC_count_grouped.csv",row.names = F,quote = F)

unique(EC_c$EC) %>% length()

```

## KEGG_Pathway (one gene to many )
```{r}
count_colname<-c("Query","KEGG_Pathway",colnames(Nmapped_count) )
KEGG_Pathway <- annot_count[,count_colname]
KEGG_Pathway <- KEGG_Pathway[KEGG_Pathway$KEGG_Pathway != "-",]
KEGG_Pathway <- KEGG_Pathway[,-1]


KEGG_Pathway_extended <- c()
file.remove("Genecatalog/annotations/KEGG_Pathway_count.tsv")


write.table(colnames(KEGG_Pathway) %>% t(),file = "Genecatalog/annotations/KEGG_Pathway_count.tsv",append = T,sep="\t",quote = F,row.names = F,col.names = F)

for (i in 1 : dim(KEGG_Pathway)[1]){
  i=1
  id=KEGG_Pathway[i,]
  if (grepl(",",id$KEGG_Pathway)){
    KEGG_Pathways <- strsplit(id$KEGG_Pathway,split = ",")[[1]]
    for(n in 1:length(KEGG_Pathways)){
      n=1
      tmp <- data.frame(KEGG_Pathways[n],id[-1] )
      colnames(tmp)[1] <- "KEGG_Pathway"
      KEGG_Pathway_extended <- bind_rows(KEGG_Pathway_extended,tmp )
    }
  }else{
  KEGG_Pathway_extended <- bind_rows(KEGG_Pathway_extended,id)
  }
  
  if(i %% 500 == 0){
    
    #print(i)
    
    write.table(KEGG_Pathway_extended,file = "Genecatalog/annotations/KEGG_Pathway_count.tsv",append = T,sep="\t",quote = F,row.names = F,col.names = F)
    KEGG_Pathway_extended <- c()
    
    }
}


# read table
KEGG_Pathway_c <- read.delim2("Genecatalog/annotations/KEGG_Pathway_count.tsv")
KEGG_Pathway_c$KEGG_Pathway <- gsub("KEGG_Pathway:","",KEGG_Pathway_c$KEGG_Pathway)

KEGG_Pathway_c2 <- KEGG_Pathway_c %>% arrange(KEGG_Pathway) %>% 
  group_by(KEGG_Pathway) %>%
  summarise_all(.funs = sum)

write.csv(KEGG_Pathway_c2,file = "Genecatalog/annotations/KEGG_Pathway_count_grouped.csv",row.names = F,quote = F)




```



## PFAMs (one gene to many )
```{r}
count_colname<-c("Query","PFAMs",colnames(Nmapped_count) )
PFAMs <- annot_count[,count_colname]
PFAMs <- PFAMs[PFAMs$PFAMs != "-",]
PFAMs <- PFAMs[,-1]


PFAMs_extended <- c()
file.remove("Genecatalog/annotations/PFAMs_count.tsv")


write.table(colnames(PFAMs) %>% t(),file = "Genecatalog/annotations/PFAMs_count.tsv",append = T,sep="\t",quote = F,row.names = F,col.names = F)

for (i in 1 : dim(PFAMs)[1]){
  id=PFAMs[i,]
  if (grepl(",",id$PFAMs)){
    PFAMss <- strsplit(id$PFAMs,split = ",")[[1]]
    for(n in 1:length(PFAMss)){
      tmp <- data.frame(PFAMss[n],id[-1] )
      colnames(tmp)[1] <- "PFAMs"
      PFAMs_extended <- bind_rows(PFAMs_extended,tmp )
    }
  }else{
  PFAMs_extended <- bind_rows(PFAMs_extended,id)
  }
  
  if(i %% 500 == 0){
    
    #print(i)
    
    write.table(PFAMs_extended,file = "Genecatalog/annotations/PFAMs_count.tsv",append = T,sep="\t",quote = F,row.names = F,col.names = F)
    PFAMs_extended <- c()
    
    }
}


# read table
PFAMs_c <- read.delim2("Genecatalog/annotations/PFAMs_count.tsv")
PFAMs_c$PFAMs <- gsub("PFAMs:","",PFAMs_c$PFAMs)

PFAMs_c2 <- PFAMs_c %>% arrange(PFAMs) %>% 
  group_by(PFAMs) %>%
  summarise_all(.funs = sum)

write.csv(PFAMs_c2,file = "Genecatalog/annotations/PFAMs_count_grouped.csv",row.names = F,quote = F)

unique(PFAMs_c$PFAMs) %>% length()


```







